\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2018-02-04 document/XSGRB_sample.tex
Last bursts and edits

\abstract{In this work we present spectra of all $\gamma$-ray burst (GRB)
	afterglows that have been promptly observed with the X-shooter spectrograph up
	to \termdate. This totals spectroscopic observations of 102 individual bursts
	observed with 48 hours of the GRB trigger. Redshifts have been measured for 96
	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
	of observational selection criteria that minimize biases with regards to
	intrinsic properties of the GRBs, the follow-up effort has been focused on
	producing a homogeneous sample of 91 afterglow spectra for GRBs discovered by
	the {\em Swift} satellite. We here provide a public release of all the reduced
	spectra, including continuum estimates and telluric absorption corrections. For
	completeness, we also provide reductions for the 18 late-time observations of the
	underlying host galaxies. We provide an assessment of the degree of completeness
	with respect to the parent GRB population, in terms of the X-ray properties of
	the bursts in the sample and find that the sample presented here is
	representative of the full \textit{Swift} sample. For all bursts (41) for which
	it is possible, we provide a measurement of the neutral hydrogen column density,
	increasing the total number of published H{\sc I} column density measurements by
	$\sim$ 50 per cent. %	We
	% constrain the
	% fraction
	%	of dark bursts to be less than 28 per cent based on the detection of an
	% optical
	%	afterglow and tentatively confirm previous results that larger optical
	% darkness
	%	is correlated with increased X-ray absorption.
	This dataset provides a unique resource to study the ISM across cosmic time,
	from the local progenitor surroundings to the galaxy-scale conditions.}

@@

\begin{figure*}
	\centerline{\includegraphics{figures/skymap.pdf}} \caption{Mollweide projection
	in galactic coordinates of the full sky showing the positions on the sky of the
	bursts presented in this work. The equator is the Galactic plane. The empty
	stars/circles are the positions of all the 1266 \textit{Swift} bursts detected
	until \termdate. Stars indicate bursts with measured redshifts and circles
	indicate those without. Blue stars indicate the position of the 98 bursts
	fulfilling the sample criteria specified in Sect. \ref{samplecrit} that have a
	measured redshift. Red outlines are added to the 77 GRBs that enter our
	statistical sample with both X-shooter spectroscopy and a measured redshift.
	Red outlines of empty stars represent bursts which is observed with X-shooter
	and has a measured redshift, but is not a part of our statistical sample. The
	blue dots show the positions of the 63 GRBs of our statistical sample that lack
	redshift measurements. A red outline is added around the 8 bursts in our
	statistical sample which was observed with X-shooter, but did not yield a
	redshift measurement. A single empty circle with a red outline indicate a burst
	outside the statistical sample that was followed up with X-shooter, but without
	a redshift measurement. The different samples are compared in Sect.
	\ref{results}. The background shows the dust maps presented in
	\citet{Schlegel1998}. Please note that we removed the background where the
	sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a white
	color visible at the equator. The grey-scale bar below indicates the value of
	$A_V$ on the plot. The dotted lines indicate intervals of 30$^\circ$ in
	longitude and latitude.} \label{fig:skymap}
\end{figure*}

@@

\begin{figure}[!ht]
	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
	horizontal axis is in units pixels, the top axis in arc seconds. The red
	datapoints show a subsample of measurements obtained for NIR spectra. The grey
	line shows a linear fit to the VIS datapoints. } \label{fig:res}
\end{figure}

@@

To this end we use observations of telluric standard stars that are taken with
identical instrument settings as our afterglow spectra, usually just after the
science data, as part of the ESO X-shooter calibration plan. These spectra have
been reduced together with the afterglow spectra, using identical pipeline
settings with the same version of the pipeline. First we fit a Gaussian function
in the spatial direction of the trace of the standard star at 792 nm (i.e. in
the VIS arm). After this, we fit a series of  20 telluric absorption lines in
the telluric standard star spectra with Gaussians, taking care to select
transitions that are not almost-resolved multiples, should be intrinsically
unresolved, and are in areas with well defined continuum flux. We pick 34
telluric standard stars spanning a range of DIMM seeing values, with the
majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
\ref{fig:res}). We use this linear relation as a way to estimate the spectral
resolution for medium to low signal to noise afterglow spectra in the VIS arm.
To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
arm spectra of a subset of 19 sources used for the VIS arm above, and find that
the resulting distribution is consistent with a simple scaling of the VIS arm
relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
filling, sources as given on the ESO instrument website. The UVB arm contains no
suitable absorption lines to use, and we therefore use a scaled value as in the
NIR arm. This simple analysis gives a sufficiently accurate estimate for the
analysis of the low signal-to-noise science spectra.

@@

We compare the point estimates for the redshift distributions of previous
complete samples of GRBs in Tab. \ref{tab:redshift_comparison}. We see that when
we compare to other complete samples, the XS-GRB presented here has the lowest
average redshift. However, the other samples also exhibit a large spread in the
redshift distributions. A 2-sided KS test confirms that there is only mild
evidence against the Fynbo09 (p-value = 0.05) and TOUGH (p-value = 0.05) samples
being drawn from the same parent sample, while XS-GRB is consistent with the
SHOALS (p-value = 0.08) and BAT6 (p-value = 0.86) samples.

@@

In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{HI, X}}$
distribution with the one presented in \citet{Fynbo2009}. We take the
$N_{\mathrm{HI,X}}$ values from the
\href{http://www.swift.ac.uk/xrt_spectra}{XRT spectral fits} \citep{Evans2009}.
The values from \citet{Fynbo2009} have been treated as detections, meaning that
we artificially bias the distribution towards higher $\beta_{OX}$-values. The
two distributions exhibit a large degree of overlap. We cannot confirm the
result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher
\nhx. Specifically, for bursts with measured redshift either from the afterglow
or the host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$
we find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
\nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
within the error intervals. A 2-sided KS test fails to reject the null
hypothesis that they are drawn from the same distribution with $p = 0.43$,
meaning that there is no strong evidence for a discrepancy. A Kendall's $\tau$
test suggest a non-significant, very low degree of negative correlation ($\Gamma
= -0.15$ at a p-value = 0.10).

@@

\begin{figure*}[!h]
	\centering
	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
	\caption{Measurements of the hydrogen column-densities for all bursts with a
		clear Lyman alpha absorption system. In solid black is shown the spectrum with
		black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
		flux density. The solid red line is the absorption of column density equal to
		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
		with dashed lines.}
	\label{fig:HI1}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI2}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI3}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI4}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI5}
\end{figure*}
\clearpage

@@

The data presented here also formed the basis of GCN 10620 \citep{GCN10620} and GCN 10631 \citep{GCN10631} and is published
in \citet{DeUgartePostigo2011}. The burst have been followed up in three epochs
of observations, 0.4, 1.4, and 2.4 days after the burst, each lasting 4.8 ks.
The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous
detection of emission features belonging to nebular lines, including \hi, \oii,
\oiii, \neiii, \nii, \sii, \siii, and \hei~as well as absorption features due to
the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all
at a consistent redshift. Temporal evolution of the fine structure lines
belonging to \feii$^*$ is found between the epochs.

@@

The data presented here has previously been published in \citet{Michaowski2016}.
2.4ks spectroscopic integration of the host galaxy, obtained long after the
burst had faded, contains bright emission lines filling the entire slit on top
of a broad, underlying stellar continuum. We identify emission lines from \oii,
\hd, \hg, \hb, \oiii, \nii, \hb, \sii, \ariii, and \sii, all at $z=0.013$.
Significant velocity structure of the lines across the spatial direction of the
slit indicates a large degree of coherent motion relative to the line-of-sight.

@@

\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)} \label{120805}

A separate reduction of this burst has been published in \citet{Kruhler2015},
but is not used otherwise. Starting 9 days after the burst trigger, this is host
observation and does not contain any afterglow continuum. In 3.6 ks integration
time, we detect a faint continuum visible from 450 nm and all the way through
2100 nm, in contrast to what is found previously. The continuum from 4500 - 600
nm is detected at very low significance. If the drop at 450 nm is the Lyman
limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim 3.9$. The
absence of nebular lines if due to \oii~falling in a telluric absorption band
and the rest being shifted out of the wavelength coverage.

@@

@@

\subsection{GRB~121209A (z = 2.1)} \label{121209}

This data is published in \citet{Kruhler2015}, but is not used elsewhere. This
short spectroscopic integration of 1.2ks reveals a very faint trace visible in
the visual arm after binning. The spectrum is taken at high airmass and
observations were discontinued due to the hardware limit of the telescope. No
emission lines are detected in the spectrum. The redshift is adopted from
\citet{Kruhler2015}.

@@

This burst is the first sGRB observed with a potential associated
kilonova \citep{Tanvir2013a, Berger2013a}.  GCN 14757 \citep{GCN14757} was based
on this spectrum, but it is not published elsewhere. Starting 8.2 hours after
the \swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
Continuum is clearly detected across all arms from both host and afterglow and
superposed are both absorption (\cahk~and \mgii) and emission lines (\oii, \hb,
\oiii, \ha, and \sii), all at a consistent redshift of $z = 0.356$, which is the
redshift of the GRB.

@@

\subsection{GRB~130615A (z $\sim 3$)} \label{130615}

This spectrum has not previously been published. Starting only 45 minutes after
the BAT trigger, $2\times600$ s spectroscopic integration carried on into the
beginning twilight. Observed at very high airmass with a quickly varying
background, a faint afterglow trace is visible across all arms of X-shooter,
down to ~ 4800 \AA, confirming the approximate redshift suggested in GCN 14898
\citep{GCN14898}.

@@

This spectrum has already been used in GCN 15250 \citep{GCN15250} and is
additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
Observations of this burst began with X-shooter 3.5 hours after the
\swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, as well as those
from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).

@@

These data formed the basis for GCN 15900 \citep{GCN15900} and is additionally
published in \citet{Kruhler2015}. Spectroscopic follow-up began 9 hours after
the BAT trigger and lasted for $12\times600$ s. A low signal-to-noise, spatially
extended continuum is visible across the entire spectral coverage of X-shooter.
The GRB afterglow continuum is visible at moderate signal-to-noise on top of the
underlying host continuum. The host exhibits usual nebular emission from \oii,
\hb, \oiii, \ha, \nii, and \sii which puts it at $z = 1.416$. Supporting this as
the redshift of the GRB are absorption features from the \mgii-doublet in the
GRB afterglow continuum.

@@

This gross content of this data has previously been issued in GCN 17832
\citep{GCN17832}, but is not published. Starting $> 1$ day after the GRB
trigger, $4 \times 600$ s spectroscopic integration securely allows us to
measure the redshift of the host. A continuum is detected all the way through
the spectral coverage with multiple emission lines superposed. It is not clear
to what degree the GRB afterglow contributes to the continuum. We identify the
emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and \sii~- all at $z =
0.256$. Due to the spatial proximity of this burst, it is a candidate for SN
follow-up and indeed there are indications of re-brightening at the burst
position (GCN 17903 \citep{GCN17903}).

@@

The gross structure of these spectra have previously been issued in GCN 19192
\citep{GCN19192}. Spectroscopic integration began 13 hours post GRB trigger and
lasted for $4 \times 1200$ s. In the data we see a moderate signal-to-noise
continuum across the entire spectral coverage of X-shooter. Clearly visible on
top of the continuum are emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
\nii, and \sii~at a consistent redshift of $z = 0.726$. It is unclear how much
of the continuum is due to the host and how much is from the GRB afterglow. At
the same redshift, we see the tentative detection of \feii-absorption,
supporting the redshift measurement and suggesting a significant afterglow
contribution to the continuum.

@@

These spectra were already used to form the basis for GCN 19350
\citep{GCN19350}, but are not published anywhere else. Spectroscopic integration
started after 7.2 hours and lasted for $4 \times 1200$ s. A low signal-to-noise
continuum is detected across the spectral coverage of X-shooter, but no clear
absorption lines are easily discernible. It is therefore unclear how much of the
continuum is due to the host and how much is due to the GRB. Superposed on the
continuum are two sets of emission lines, offset spatially, but at very similar
redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and \sii~in both
systems -- all at $z = 0.555$. As per GCN 19350 \citep{GCN19350}, the GRB
probably occurred in an interacting pair of galaxies.

@@

The data presented here also formed the basis of GCN 19773 \citep{GCN19773}, and
is published in \citet{Heintz2017b}. Observations started 22.37 hours after the
BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the
entire spectral coverage of X-shooter and absorption lines from \mgi, \mgii,
\feii~and \alii~are found at $z = 0.736$. At the same redshift, emission lines
from \oii, \oiii, \ha, \hb, \hg, \nii, \sii, \siii~are found. A second epoch,
lasting 3.6ks, is obtained after the afterglow has faded, confirming the
emission line detections. The host galaxy is found to have a roughly solar
metallicity and is among the most luminous GRB hosts at $z < 1$.

@@

The content and scope of these spectra has previously been issued in GCN 19971
\citep{GCN19971}. $4 \times 600$ s spectroscopic integration was obtained of
this potentially short GRB (GCN 19974\citep{GCN19974}), starting 6.1 hours after
the GRB trigger. The main source in the slit is likely of stellar origin and
thus a foreground object unrelated to the GRB. Superposed, but slightly offset,
from the stellar source are emission lines from \oii, \hb, and \ha~at a
consistent redshift of $z = 0.891$. As stated in GCN 19971 \citep{GCN19971}, the
association between the GRB and the source at $z = 0.891$ is uncertain. An
alternative host association is suggestion in GCN 19975 \citep{GCN19975} just
outside the XRT error circle. The fading of the source in GCN 19975
\citep{GCN19975} is, however, not confirmed (Kruehler, private communication)
and thus a clear localization unclear. Due to the proximity of the $z =
0.891$-host suggested here to the XRT localization, we consider this redshift as
the most likely redshift of the GRB.

>>>>

\abstract{In this work we present spectra of all $\gamma$-ray burst (GRB)
	afterglows that have been promptly observed with the X-shooter spectrograph up
	to \termdate. This totals spectroscopic observations of 103 individual bursts
	observed with 48 hours of the GRB trigger. Redshifts have been measured for 97
	per cent of these, covering a redshift range from 0.059 to 7.84. Based on a set
	of observational selection criteria that minimize biases with regards to
	intrinsic properties of the GRBs, the follow-up effort has been focused on
	producing a homogeneous sample of 93 afterglow spectra for GRBs discovered by
	the {\em Swift} satellite. We here provide a public release of all the reduced
	spectra, including continuum estimates and telluric absorption corrections. For
	completeness, we also provide reductions for the 18 late-time observations of the
	underlying host galaxies. We provide an assessment of the degree of completeness
	with respect to the parent GRB population, in terms of the X-ray properties of
	the bursts in the sample and find that the sample presented here is
	representative of the full \textit{Swift} sample. For all bursts (42) for which
	it is possible, we provide a measurement of the neutral hydrogen column density,
	increasing the total number of published H{\sc I} column density measurements by
	$\sim$ 50 per cent. %	We
	% constrain the
	% fraction
	%	of dark bursts to be less than 28 per cent based on the detection of an
	% optical
	%	afterglow and tentatively confirm previous results that larger optical
	% darkness
	%	is correlated with increased X-ray absorption.
	This dataset provides a unique resource to study the ISM across cosmic time,
	from the local progenitor surroundings to the galaxy-scale conditions.}

@@

\begin{figure*}
	\centerline{\includegraphics{figures/skymap.pdf}} \caption{Mollweide projection
	in galactic coordinates of the full sky showing the positions on the sky of the
	bursts presented in this work. The equator is the Galactic plane. The empty
	stars/circles are the positions of all the 1266 \textit{Swift} bursts detected
	until \termdate. Stars indicate bursts with measured redshifts and circles
	indicate those without. Blue stars indicate the position of the 104 bursts
	fulfilling the sample criteria specified in Sect. \ref{samplecrit} that have a
	measured redshift. Red outlines are added to the 84 GRBs that enter our
	statistical sample with both X-shooter spectroscopy and a measured redshift.
	Red outlines of empty stars represent bursts which is observed with X-shooter
	and has a measured redshift, but is not a part of our statistical sample. The
	blue dots show the positions of the 61 GRBs of our statistical sample that lack
	redshift measurements. A red outline is added around the six bursts in our
	statistical sample which was observed with X-shooter, but did not yield a
	redshift measurement. Two empty circles with a red outlines indicate bursts
	outside the statistical sample that was followed up with X-shooter, but without
	a redshift measurement. The different samples are compared in Sect.
	\ref{results}. The background shows the dust maps presented in
	\citet{Schlegel1998}. Please note that we removed the background where the
	sample criteria is violated ($A_V > 0.5$ mag) and replaced it with a white
	color visible at the equator. The grey-scale bar below indicates the value of
	$A_V$ on the plot. The dotted lines indicate intervals of 30$^\circ$ in
	longitude and latitude.} \label{fig:skymap}
\end{figure*}

@@

%\begin{figure}[!ht]
%	\centerline{\includegraphics[width=8cm]{figures/resolution_paper.eps}}
%\caption{Green datapoints show the FWHM (km/s) of Gaussian fits to unresolved
%	telluric absorption lines in the VIS spectra, as a function of the FWHM of a
%	Gaussian fit onto the trace in the spatial direction at 792 nm. The lower
%	horizontal axis is in units pixels, the top axis in arc seconds. The red
%	datapoints show a subsample of measurements obtained for NIR spectra. The grey
%	line shows a linear fit to the VIS datapoints. } \label{fig:res}
%\end{figure}

\begin{figure}[!ht]
	\centerline{\includegraphics[width=8cm]{figures/resolutions}} \caption{Blue
	datapoints show the FWHM (km/s) of Gaussian fits to unresolved telluric
	absorption lines in the VIS spectra, as a function of the airmass corrected
	ambient FWHM. The red datapoints show a subsample of measurements obtained for
	NIR spectra. The colored lines show the corresponding linear fits to the
	datapoints. The colored bounds contain 68 per cent of the best-fit probability
	mass. } \label{fig:res}
\end{figure}

@@

To this end we use observations of telluric standard stars that are taken with
identical instrument settings as our afterglow spectra, usually just after the
science data, as part of the ESO X-shooter calibration plan. These spectra have
been reduced together with the afterglow spectra, using identical pipeline
settings with the same version of the pipeline.

%First we fit a Gaussian function
%in the spatial direction of the trace of the standard star at 792 nm (i.e. in
%the VIS arm). After this, we fit a series of  20 telluric absorption lines in
%the telluric standard star spectra with Gaussians, taking care to select
%transitions that are not almost-resolved multiples, should be intrinsically
%unresolved, and are in areas with well defined continuum flux. We pick 34
%telluric standard stars spanning a range of DIMM seeing values, with the
%majority between $0.5-1.5 $\arcsec. The resulting distribution of spectral FWHM
%(km/s) as a function of spatial FWHM at 792 nm is fairly well described by a
%linear relation $a + b*x$, with $x$ the spatial FWHM in pixels (with 0.15 arc
%sec per pixel),  $a= 21.4 \pm 1.3$ km/s, $b=1.4 \pm 0.2$ (see Fig.
%\ref{fig:res}). We use this linear relation as a way to estimate the spectral
%resolution for medium to low signal to noise afterglow spectra in the VIS arm.
%To extend this to the UVB and NIR arm, we measured a series of lines in the NIR
%arm spectra of a subset of 19 sources used for the VIS arm above, and find that
%the resulting distribution is consistent with a simple scaling of the VIS arm
%relation by the ratio of resolutions of the NIR and VIS arm for unresolved, slit
%filling, sources as given on the ESO instrument website. The UVB arm contains no
%suitable absorption lines to use, and we therefore use a scaled value as in the
%NIR arm. This simple analysis gives a sufficiently accurate estimate for the
%analysis of the low signal-to-noise science spectra.

%fit parameter 1-sigma error
%———————————-
%6.96733427249 +- 1.35492333728
%36.1113577532 +- 1.36818568362
%fit parameter 1-sigma error
%———————————-
%3.64773146238 +- 0.563740522948
%24.2308362925 +- 0.792580030602

@@

We compare the point estimates for the redshift distributions of previous
complete samples of GRBs in Tab. \ref{tab:redshift_comparison}. We see that when
we compare to other complete samples, the XS-GRB presented here has the lowest
average redshift. However, the other samples also exhibit a large spread in the
redshift distributions. A 2-sided KS test reveals that the XS-GRB sample is consistent with
being drawn from the same parent sample with the following p-values: SHOALS (p-value = 0.13), BAT6 (p-value = 0.95), Fynbo09 (p-value = 0.08) and TOUGH (p-value = 0.09).

@@

In Fig. \ref{fig:betaOX} we compare the $\beta_{OX}$ - $N_{\mathrm{HI, X}}$
distribution with the one presented in \citet{Fynbo2009}. We take the
$N_{\mathrm{HI,X}}$ values from the
\href{http://www.swift.ac.uk/xrt_spectra}{XRT spectral fits} \citep{Evans2009}.
The values from \citet{Fynbo2009} have been treated as detections, meaning that
we artificially bias the distribution towards higher $\beta_{OX}$-values. The
two distributions exhibit a large degree of overlap. We confirm the
result by \citet{Fynbo2009}, that dark bursts, $\beta_{OX} < 0.5$, have higher
\nhx. Specifically, for bursts with measured redshift either from the afterglow
or the host galaxy, we find the following: For bursts with $\beta_{OX} \geq 0.5$
we find \nhx~$ = 21.4_{-1.0}^{+0.7}$, whereas for $\beta_{OX} < 0.5$ we find
\nhx~$21.8_{-0.9}^{+0.5}$ where 68 per cent of the probability mass is contained
within the error intervals. A 2-sided KS test fails to reject the null
hypothesis that they are drawn from the same distribution with $p = 0.11$,
meaning that there is no strong evidence for a discrepancy. A Kendall's $\tau$
test, however, suggest a statistically significant, low degree of negative
correlation ($\Gamma = -0.21$ at a p-value = 0.01).

@@

\begin{figure*}[!h]
	\centering
	\includegraphics[page=1, width=16cm]{figures/HI_measurements.pdf}
	\caption{Measurements of the hydrogen column-densities for all bursts with a
		clear Lyman alpha absorption system. In solid black is shown the spectrum with
		black dotted giving the corresponding 1-$\sigma$ error. Black dashed shows zero
		flux density. The solid red line is the absorption of column density equal to
		the value presented in Tab. \ref{tab:HI} with the 1-$\sigma$ interval shown
		with dashed lines.}
	\label{fig:HI1}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=2, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI2}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=3, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI3}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=4, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI4}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=5, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI5}
\end{figure*}
\clearpage
\begin{figure*}[!h]
	\centering
	\includegraphics[page=6, width=16cm]{figures/HI_measurements.pdf}
	\caption*{Fig. \ref{fig:HI1}. continued.}
	\label{fig:HI6}
\end{figure*}
\clearpage

@@

The data presented here also formed the basis of GCN 10620 \citep{GCN10620} and GCN 10631 \citep{GCN10631} and is published
in \citet{DeUgartePostigo2011}. The burst have been followed up in three epochs
of observations, 0.4, 1.4, and 2.4 days after the burst, each lasting 4.8 ks.
The unambiguous redshift of the host, $z=0.624$, is found from the simultaneous
detection of emission features belonging to nebular lines, including \hi, \oii,
\oiii, \neiii, \nii, [\sii], [\siii], and \hei~as well as absorption features due to
the presence of \znii, \crii, \feii, \mnii, \mgii, \mgi, \tiii, and \caii, all
at a consistent redshift. Temporal evolution of the fine structure lines
belonging to \feii$^*$ is found between the epochs.

@@

The data presented here has previously been published in \citet{Michaowski2016}.
2.4ks spectroscopic integration of the host galaxy, obtained long after the
burst had faded, contains bright emission lines filling the entire slit on top
of a broad, underlying stellar continuum. We identify emission lines from \oii,
\hd, \hg, \hb, \oiii, \nii, \hb, [\sii], \ariii, and [\siii], all at $z=0.013$.
Significant velocity structure of the lines across the spatial direction of the
slit indicates a large degree of coherent motion relative to the line-of-sight.

@@

\subsection{GRB~120805A (z = 3.9 NEW!!!)} \label{120805}

A separate reduction of this burst has been published in \citet{Kruhler2015},
but is not used otherwise. Starting 9 days after the burst trigger, this is host
observation and does not contain any afterglow continuum. In 3.6 ks integration
time, we detect a faint continuum visible from 450 nm and all the way through
2100 nm, in contrast to what is found previously. The continuum from 450 - 600
nm is detected at very low significance. If the drop at 450 nm is the Lyman
limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim 3.9$. The
absence of nebular lines is due to \oii~falling in a telluric absorption band
and the rest of the nebular lines being shifted out of the wavelength coverage.

@@

This spectrum has previously been published in \citet{Tanvir2017a}. Starting
18.5 hours after the BAT trigger, the final spectrum is based on $2 \times 4
\times 1200$ s spectroscopic integration for a total of 160 minutes. Nothing is
immediately visible in the 2D-spectra, however after severe binning in the
dispersion direction, a faint trace shows up in the NIR arm, suggesting a very
high redshift. We here adopt the redshift ($z = 7.84$) suggested by
\citet{Tanvir2017a}.

@@

%\subsection{GRB~121209A (z = 2.1)} \label{121209}
%
%This data is published in \citet{Kruhler2015}, but is not used elsewhere. This
%short spectroscopic integration of 1.2ks reveals a very faint trace visible in
%the visual arm after binning. The spectrum is taken at high airmass and
%observations were discontinued due to the hardware limit of the telescope. No
%emission lines are detected in the spectrum. The redshift is adopted from
%\citet{Kruhler2015}.

@@

This burst is the first sGRB observed with a potential associated
kilonova \citep{Tanvir2013a, Berger2013a}.  GCN 14757 \citep{GCN14757} was based
on this spectrum, but it is not published elsewhere. Starting 8.2 hours after
the \swift~trigger, a total of 2.4 ks spectroscopic integration was obtained.
Continuum is clearly detected across all arms from both host and afterglow and
superposed are both absorption (\cahk~and \mgii) and emission lines (\oii, \hb,
\oiii, \ha, and [\sii]), all at a consistent redshift of $z = 0.356$, which is the
redshift of the GRB.

@@

\subsection{GRB~130615A (z = 2.9)} \label{130615}

This spectrum has not previously been published. Starting only 45 minutes after
the BAT trigger, $2\times600$ s spectroscopic integration carried on into the
beginning twilight. Observed at very high airmass with a quickly varying
background, a faint afterglow trace is visible across all arms of X-shooter,
down to ~ 4800 \AA, which if interpreted as the break due to \lya suggests $z =
2.9$. This supports the approximate redshift suggested in GCN 14898
\citep{GCN14898}.

@@

This spectrum has already been used in GCN 15250 \citep{GCN15250} and is
additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
Observations of this burst began with X-shooter 3.5 hours after the
\swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
\hb, \oii, \ha, \nii, [\sii]) are well detected at $z = 0.347$, as well as those
from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).

@@

These data formed the basis for GCN 15900 \citep{GCN15900} and is additionally
published in \citet{Kruhler2015}. Spectroscopic follow-up began 9 hours after
the BAT trigger and lasted for $12\times600$ s. A low signal-to-noise, spatially
extended continuum is visible across the entire spectral coverage of X-shooter.
The GRB afterglow continuum is visible at moderate signal-to-noise on top of the
underlying host continuum. The host exhibits usual nebular emission from \oii,
\hb, \oiii, \ha, \nii, and [\sii] which puts it at $z = 1.416$. Supporting this as
the redshift of the GRB are absorption features from the \mgii-doublet in the
GRB afterglow continuum.

@@

This gross content of this data has previously been issued in GCN 17832
\citep{GCN17832}, but is not published. Starting $> 1$ day after the GRB
trigger, $4 \times 600$ s spectroscopic integration securely allows us to
measure the redshift of the host. A continuum is detected all the way through
the spectral coverage with multiple emission lines superposed. It is not clear
to what degree the GRB afterglow contributes to the continuum. We identify the
emission lines as \oii, \hb, \oiii, \ha, \nii, \ha, and [\sii]~- all at $z =
0.256$. Due to the spatial proximity of this burst, it is a candidate for SN
follow-up and indeed there are indications of re-brightening at the burst
position (GCN 17903 \citep{GCN17903}).

@@

The gross structure of these spectra have previously been issued in GCN 19192
\citep{GCN19192}. Spectroscopic integration began 13 hours post GRB trigger and
lasted for $4 \times 1200$ s. In the data we see a moderate signal-to-noise
continuum across the entire spectral coverage of X-shooter. Clearly visible on
top of the continuum are emission lines from \oii, \hd, \hg, \hb, \oiii, \ha,
\nii, and [\sii]~at a consistent redshift of $z = 0.726$. It is unclear how much
of the continuum is due to the host and how much is from the GRB afterglow. At
the same redshift, we see the tentative detection of \feii-absorption,
supporting the redshift measurement and suggesting a significant afterglow
contribution to the continuum.

@@

These spectra were already used to form the basis for GCN 19350
\citep{GCN19350}, but are not published anywhere else. Spectroscopic integration
started after 7.2 hours and lasted for $4 \times 1200$ s. A low signal-to-noise
continuum is detected across the spectral coverage of X-shooter, but no clear
absorption lines are easily discernible. It is therefore unclear how much of the
continuum is due to the host and how much is due to the GRB. Superposed on the
continuum are two sets of emission lines, offset spatially, but at very similar
redshifts. We identify lines from \oii, \hb, \oiii, \ha, \nii, and [\sii]~in both
systems -- all at $z = 0.555$. As per GCN 19350 \citep{GCN19350}, the GRB
probably occurred in an interacting pair of galaxies.

@@

The data presented here also formed the basis of GCN 19773 \citep{GCN19773}, and
is published in \citet{Heintz2017b}. Observations started 22.37 hours after the
BAT trigger and lasted for 2.4ks. The afterglow continuum is detected across the
entire spectral coverage of X-shooter and absorption lines from \mgi, \mgii,
\feii~and \alii~are found at $z = 0.736$. At the same redshift, emission lines
from \oii, \oiii, \ha, \hb, \hg, \nii, [\sii], [\siii]~are found. A second epoch,
lasting 3.6ks, is obtained after the afterglow has faded, confirming the
emission line detections. The host galaxy is found to have a roughly solar
metallicity and is among the most luminous GRB hosts at $z < 1$.

@@

The content and scope of these spectra has previously been issued in GCN 19971
\citep{GCN19971}. $4 \times 600$ s spectroscopic integration was obtained of
this potentially short GRB \citep[GCN 19974;][]{GCN19974}, starting 6.1 hours
after the GRB trigger. The main source in the slit is likely of stellar origin
and thus a foreground object unrelated to the GRB. Superposed, but slightly
offset, from the stellar source are emission lines from \oii, \hb, and \ha~at a
consistent redshift of $z = 0.891$. As stated in GCN 19971 \citep{GCN19971}, the
association between the GRB and the source at $z = 0.891$ is uncertain. An
alternative host association is suggestion in GCN 19975 \citep{GCN19975} just
outside the XRT error circle. The fading of the source in GCN 19975
\citep{GCN19975} is, however, not confirmed (Kruehler, private communication)
and thus a clear localization unclear. Due to the proximity of the $z =
0.891$-host suggested here to the XRT localization, we consider this redshift as
the most likely redshift of the GRB.
